---
title: "3_charts"
output: html_document
---
installing libraries I might need
```{r}
library(tidyverse)
library(scales)
library(treemapify)
library(RColorBrewer)
library(svglite)
```

calling spotify data

```{r}
spotify_data <- read.csv("Most_Streamed_Spotify _Songs_2023.csv")
```

here's the first thing I wanna chart: 
top ten songs in most spotify playlists
```{r}
spotify_bar_chart_data <- spotify_data %>% top_n(10, in_spotify_playlists)
```

now organizing in descending order
```{r}
spotify_bar_chart_data %>% ggplot(aes(x = reorder(track_name,desc(in_spotify_playlists)), y = in_spotify_playlists)) +
  geom_col()
```
adding more specifics:
```{r}
spotify_bar_chart_data %>% ggplot(aes(x = reorder(track_name,desc(in_spotify_playlists)), y = in_spotify_playlists)) +
  geom_col(fill = "red")+
  scale_y_continuous(labels = scales::label_number()) +
  scale_x_discrete(guide = guide_axis(angle = 45)) +
  labs(title = "Songs on the most Spotify playlists in 2023", subtitle = "Daft Punk's 2013 hit 'Get Lucky' topped the list on nearly 53k playlists", caption = "Source: Spotify via Kaggle \n by Sarah Meklir")+
  theme_minimal()+
  theme(axis.title = element_blank())
```

Awesome! On to the next chart. This one will be a line chart looking at streams based on the year they were released. 
```{r}
spotify_area_data <- spotify_data %>% top_n(10, streams) %>% arrange(desc(released_year))

spotify_area_data %>% ggplot(aes(x = released_year, y = streams)) +
  geom_line(alpha = 0.5, fill="darkred") +   
  labs(title = "Studying Spotify Streams by Release Year", subtitle = "2023 Spotify stream spikes in songs released in 2017 and 2019", caption = "Source: Spotify via Kaggle \n by Sarah Meklir", x = "Year", y = "Count")
```

I mean sure! That's not really what I had in mind, but I've tried and I'm not sure how to get it. On to scatter plots!

Going to compare bpm and danceability to see if those are linked. My guess is yes.
```{r}
  ggplot(spotify_data, aes(bpm, danceability_., color = energy_.)) +
    geom_point()
```
Interesting! It looks like danceability and bpm aren't inextricably linked, and it's cool to see the energy tending to be lower at lower danceability. 

Adding chatter and headings: 
```{r}
  ggplot(spotify_data, aes(bpm, danceability_., color = energy_.)) +
    geom_point() +
labs(title = "Are BPM (beats per minute) and danceability related?", subtitle = "Examining the top-streamed songs on Spotify in 2023", caption = "Source: Spotify via Kaggle \n by Sarah Meklir", x = "BPM", y = "Danceability", color = "Energy")
```